An Efficient Forecasting Approach\\ to Reduce Boundary Effects in Real-Time Time-Frequency Analysis

### 2020-09-03 Notes/BoundEffRed.tex
update paper

Various improvements can be considered to make the algorithm more robust. In particular, we have noticed (see Fig.~\ref{fig:THO.failure}) that when the regular oscillations of the observed signal break, the forecasting step is no longer relevant, and only slows down the calculation of the TF representation. A preliminary step should then be added to the algorithm to detect signal activity, and disable the forecasting step when possible. More fundamentally, one can also consider accelerating the computational time by optimizing the forecasting step, \ie~reduce $t_{\mathrm{forecast}}$. Indeed, each new forecast requires the inversion of the matrix $\bX\bX^T$ of size $M\times M$. However, the matrix $\bX$ used to forecast $\bx_{N+H}$ at the current iteration differs from the one used at the previous iteration to forecast $\bx_{N}$ by only $H$ columns. It then seems natural to take inspiration from the work of Strobach~\cite{Strobach97square}, generalized by Badeau \etal~\cite{Badeau04sliding}, which proposes a fast algorithm for the singular value decomposition of successive data matrices taking the same form as $\bX$. Their algorithm is limited to the case where $H=1$. Developing an extension of this algorithm to variations of $H>1$ columns would then allow to efficiently update the matrix $\left(\bX\bX^T\right)^{-1}$. These avenues will be explored in our future work.

>>>>

Various improvements can be considered to make the algorithm more robust. In particular, we have noticed (see Fig.~\ref{fig:THO.failure}) that when the regular oscillations of the observed signal break, the forecasting step is no longer relevant, and only slows down the calculation of the TF representation. A preliminary step should then be added to the algorithm to detect signal activity, and disable the forecasting step when possible. More fundamentally, one can also consider accelerating the computational time by optimizing the forecasting step to improve the real-time performance of {\sf BoundEffRed}. Indeed, each new forecast requires the inversion of the matrix $\bX\bX^T$ of size $M\times M$. However, the matrix $\bX$ used to forecast $\bx_{N+H}$ at the current iteration differs from the one used at the previous iteration to forecast $\bx_{N}$ by only $H$ columns. It then seems natural to take inspiration from the work of Strobach~\cite{Strobach97square}, generalized by Badeau \etal~\cite{Badeau04sliding}, which proposes a fast algorithm for the singular value decomposition of successive data matrices taking the same form as $\bX$. Their algorithm is limited to the case where $H=1$. Developing an extension of this algorithm to variations of $H>1$ columns would then allow to efficiently update the matrix $\left(\bX\bX^T\right)^{-1}$. These avenues will be explored in our future work.
